Eclipse Che for Fair and Easy Education

### 2019-05-04 paper.tex
Raw transcription of Eclipse Che

@@

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  Stackoverflow  developersurvey [22] showed, that DevOps-engineers are not only 
amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
So why there are no courses? There are many reasons, but the most prominent is the complexity of the topic: 
To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
infrastructure  and  networking. 
Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
All these factors make a high stake for entrance, also  represented in the developersurvey [22] which states that most DevOps engineers have a decade of experience, usually in operations topics.
Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
Che is unlike a normal eclipse distribution: 
It is hosted on a server and developers get access with their browser. 
Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
The software build is run inside a docker-container, making it cloud-native by default. 
With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
The latest distributions of RedHats OpenShift are shipped with Che in default. 
Therefore, it is worth looking at this possible game-changer. 
With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.    
\section{Demonstrationen}
\label{sec:demos}
Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
Es ist kein Weierstraß-p ($\wp$) mehr.

>>>>

Figure 1

@@

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  Stackoverflow  developersurvey [22] showed, that DevOps-engineers are not only 
amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
So why there are no courses? There are many reasons, but the most prominent is the complexity of the topic: 
To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
infrastructure  and  networking. 
Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
All these factors make a high stake for entrance, also  represented in the developersurvey [22] which states that most DevOps engineers have a decade of experience, usually in operations topics.
Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
Che is unlike a normal eclipse distribution: 
It is hosted on a server and developers get access with their browser. 
Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
The software build is run inside a docker-container, making it cloud-native by default. 
With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
The latest distributions of RedHats OpenShift are shipped with Che in default. 
Therefore, it is worth looking at this possible game-changer. 
With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.

\section{Eclipse Che}
\subsection{Environment and Requirements}
Che is hosted as a server-application and is already cloudnative. 
The only requirements are that either Docker, Kubernetes or OpenShift are available for the installation.
The differences between these will shortly be summarized.

Docker \footnote{https://www.docker.com/} is a container-plattform, where a container is a standardized unit of of software including the code, dependencies and core-functionalities. 
A container runs on an OS-Virtualization, originating from LXC, has declared interfaces and can be parameterized. 
One of the core features is to connect containers for bigger projects, e.g. one 
database container and one web-server container, which are connected into a virtual network. 
Docker is mostly famous for this docker-engine, but enriches it with monitoring and logging, as well as the ability to move containers onto different machines.

Kubernetes \footnote{https://kubernetes.io/} is a container-orchestration-system based on Docker. 
Kubernetes picked up the growing problem in managing multiple docker-containers by  simplyfing  resource-management. 
Additional to deployment and monitoring of container-groups it provides auto-restart mechanisms and scaling based on (custom) metrics. 
When run on Kubernetes, Che is hosted as a single docker container in the existing platform.
The term pod, also later used in this paper, refers to a suite of connected containers. 
An example pod would be a simple two-tier web-application, where the database and webserver are two different, but connected containers.

Openshift \footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
Notable  additions  include  Jenkins,  Gitea,  Sonarqube and lately Che. 
Regarding Che there are no notable changes in handling, as Che is simply run on the build in Kubernetes.
When run on Kubernetes or Docker Che will require two additional containers for authentification with KeyCloack. 
For Openshift the standard-openshift authentification is used.  
Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations aswell as for developers working on the Che-Code. 
When run (locally) on docker, the plugin CheDir enables a portable workspace, 
which can be used on any docker based Che.
\subsection{Technology and Workflow}
A normal IDE is software on your computer which makes programming easier than doing it in a plain text-editor and commandline. 
Common tasks for an IDE are dependency management, debugging, refactoring and autocompletion.
It is important that an IDE is not required to develop – you can write .java-files in Vim, compile on the console and run a .Jar manually. 
With that in mind, we can see that the IDE is just a tool to make things (a lot) more comfortable – but you also have requirements such as the JDK, the .Jars you need in your classpath etc. as they are only utilized by your IDE.
You can consider this stack as your workspace as it contains everything required to work, with the IDE making the workspace comfortable to use.
Eclipse Che does not work like a normal IDE. 
Che consists of three main components shown in Figure 2 (TODO Figure 2).

FIGURE 2 CHES OVERVIEW

\begin{enumerate}
	\item Workspaces, including Runtimes and IDEs
	\item A browser-based IDE
	\item An administrative server
\end{enumerate}
A workspace is like your above mentioned common \textit{workspace}. 
It is a single machine, as a container, containing the project files, compilers, package managers and an IDE-Interface. 
There are two big differences to a normal development-setup:
\begin{itemize}
	\item The required items for the workspace are explicitly  declared, making  the  workspace itself a docker image
	\item The IDE is not graphical – it is a REST-API performing actions like a normal IDE, such as build, writing to files and installing packages.  
\end{itemize}
An additional distinction is about building your software.
Instead of building the .Jar in your remote-workspace, it is build inside a container in the remote-workspace.
While this sounds a bit confusing, it makes sense to separate the actual runtime from the workspace. 
This is also what happens on your machine – you compile an executable and run it separately. 
Having these runtimes inside a docker-container comes with great benefits, outlined later in this paper.
The second component is the web-based IDE, which is basically a webpage performing the required REST-calls and utility-tasks for the developer. 
It is sent by the administrative server when an authorized user accesses his workspace via browser. 
These two components can be enriched with plugins.  
Common plugins are language-extensions which support syntax-highlighting etc. or package managers with their regarding lifecycles.
The difference to a normal IDE plugin is that it has two components: 
One in the workspace-API and one in the hosted web-IDE.  
Che offers a suite of common plugins dependent on your project. 
When starting a Maven-project, the workspace is initialized with Maven and Java.
The last component is the Che-administration server. 
In the administration-server the workspaces are orchestrated and monitored. 
The authorization is also done at the admin-server and not by every workspace.
The workspaces can be split amongst multiple machines.
The runtimes do not necessarily need to be on the same machines as their workspaces. 
Instead of the web-based IDE, you can connect a desktop IDE to your workspace. 
This is done by mounting the workspaces-filesystem  in  your  desktop  IDE. 
This  will grant you access to autocompletion and other features, but you will miss Che specific helpers the web-IDE offers. 
My personal oppinion is, that mounting remote file systems for development sounds like a infinite source of problems. 
I expect this feature to be rather a proof of concept.
A single workspace can contain multiple projects, whereof every project will have its own runtime. 
Projects can either share a versioning control or be separated. 
You can build whole pods in a single workspace with the single containers being the subprojects. 
You can also compose applications inside the workspaces, meaning that the resulting 
container of a project can be the base-image of another project. 
A workspace can be shared amongst multiple users. 
Both will work on the same remote filesystem, altering files simultaneous. 
Due to changes being transmitted as HTTP-requests, the current implementation when working on the same file is a \textit{last write wins}-policy. 
For version 7 a multi-cursor file-editing will be provided, like it`s common in 
drive-documents. 
This is not to be mixed up with having multiple users work on the same project but on their own workspaces. 
The code, and the workspace itself, can commonly be shared with any versioning tool like Git.
It's mostly dependent on the team which kind of cooperation they prefer. 
As  an  interesting  fact,  Eclipse  Che  is  developed  with Eclipse Che, proving that complex cloud development is possible. 
This so-called \textit{dogfooding} usually leads to a product which puts requirements over any a-priori design or foreign interests.  
Existing components will be enhanced if working with them becomes troublesome, leaving a clean and valuable core product.   
\section{Demonstrationen}
\label{sec:demos}
Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
Es ist kein Weierstraß-p ($\wp$) mehr.
